### 2016-06-28 ansi-isolation.tex

ANSI SQL 92 standard requires \iso{Read Committted} isolation to avoid
the dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. {\sc rc}
specification is therefore a combination of these two guarantees. The
specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of {\sc rc} for highly
available replicated stores. On relational databases, however, {\sc
rc} has also come to be associated with the $\mathtt{MonotonicVis}$
guarantee.  Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$
are sufficient to reason about {\sc rc} isolation on relational stores
too. The combination of these guarantees with the {\sc sc} property of
relational stores (formalized in \S\ref{sec:store-consistency})
automatically leads to the monotonicity guarantee, which explains why
{\sc rc} comes with $\mathtt{MonotonicVis}$ on such stores regardless
of the implementation. On weakly consistent stores however,
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ do not imply
$\mathtt{MonotonicVis}$. A stronger isolation level called
\iso{Monotonic Atomic View} ({\sc mav} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat,pldi15} was proposed to
explicitly extend {\sc rc} with monotonicity on such stores. Sample
{\sc rc} executions shown in Fig.~\ref{fig:rc-executions} are also
{\sc mav}.

@@

A transaction $T_i$ is said to be snapshot-visible to a transaction
$T_j$ if either it is visible to $T_j$, or it is invisible; it is
forbidden for only a suffix (more generally, a subset) of $T_j$ to
witness $T_i$. The specification of\iso{Repeatable Read} ({\sc rr})
isolation extends {\sc mav} with the snapshot visibility guarantee.
Observe that snapshot visibility permits $T_i$ and $T_j$ to execute
and commit while being oblivious of each other. This scenario is
captured in Fig.~\ref{fig:ansi-iso-eg-rr}, where transactions $T_1$
and $T_2$, which perform conflicting writes, execute against a
snapshot of the database and commit concurrently.

\iso{Snapshot Isolation} ({\sc SI}) proscribes this possibility. If
$T_i$ and $T_j$ both write to the same shared variable, then {\sc si}
insists that either $T_i$ be visible to $T_j$, or $T_j$ be visible to
the conflicting write of $T_i$ (this is captured by the auxiliary
definition \C{SnapshotSER}). Fig.~\ref{fig:ansi-iso-eg-si} shows an
execution where {\sc si} transaction $T_1$ witnesses a snapshot of the
database that doesn't include $T_2$, but the conflicting write to $X$
in $T_2$, and the subsequent operations (due to the monotonicity
requirement of {\sc mav}) witness $T_1$. Thus, $T_2$ is aware of $T_1$
when it committed. Lastly, \iso{Serializable} isolation extends
$\mathtt{SnapshotSER}$ to also cover non-conflicting transactions that
write to variables read in the current transaction.
Fig.~\ref{fig:ansi-iso-eg-ser} shows an execution of an {\sc ser}
transacton $T_3$ and a {\sc mav transaction} $T_4$, which write to $X$
and $Y$, respectively. The snapshot witnessed by the {\sc ser}
transaction $T_3$ does not include $T_4$, but the write to $Y$ in
$T_4$, although non-conflicting, witnesses $T_1$ because $Y$ is read
by $T_1$. Visibility includes other operations of $T_4$ because of
{\sc mav}.

>>>>

The ANSI SQL 92 standard requires \iso{Read Committted} isolation to avoid
the dirty reads phenomenon, which is achieved by enforcing
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ guarantees. The {\sc rc}
specification is therefore a combination of these two guarantees. The
specification also agrees with the description and
implementation~\cite{bailishat,pldi15} of {\sc rc} for highly
available replicated stores. On relational databases, however, {\sc
rc} has also come to be associated with the $\mathtt{MonotonicVis}$
guarantee.  Nonetheless, $\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$
are sufficient to reason about {\sc rc} isolation on relational stores
too. The combination of these guarantees with the {\sc sc} property of
relational stores (formalized in \S\ref{sec:store-consistency})
automatically leads to the monotonicity guarantee, which explains why
{\sc rc} comes with $\mathtt{MonotonicVis}$ on such stores regardless
of the implementation. On weakly consistent stores however,
$\mathtt{AtomicVis}$ and $\mathtt{CommitVis}$ do not imply
$\mathtt{MonotonicVis}$. A stronger isolation level called
\iso{Monotonic Atomic View} ({\sc mav} of
Fig.~\ref{fig:ansi-isolation})~\cite{bailishat,pldi15} was proposed to
explicitly extend {\sc rc} with monotonicity on such stores. The sample
{\sc rc} executions shown in Fig.~\ref{fig:rc-executions} are also
{\sc mav}.

@@

A transaction $T_i$ is said to be snapshot-visible to a transaction
$T_j$ if either it is visible to $T_j$, or it is invisible; it is
forbidden for only a suffix (more generally, a subset) of $T_j$ to
witness $T_i$.  \iso{Repeatable Read} ({\sc rr})
isolation extends {\sc mav} with the snapshot visibility guarantee.
Observe that snapshot visibility permits $T_i$ and $T_j$ to execute
and commit while being oblivious of each other. This scenario is
captured in Fig.~\ref{fig:ansi-iso-eg-rr}, where transactions $T_1$
and $T_2$, which perform conflicting writes, execute against a
snapshot of the database and commit concurrently.

\iso{Snapshot Isolation} ({\sc SI}) proscribes this possibility. If
$T_i$ and $T_j$ both write to the same shared variable, then {\sc si}
insists that either $T_i$ be visible to $T_j$, or $T_j$ be visible to
the conflicting write of $T_i$ (this is captured by the auxiliary
definition \C{SnapshotSER}). Fig.~\ref{fig:ansi-iso-eg-si} shows an
execution where {\sc si} transaction $T_1$ witnesses a snapshot of the
database that doesn't include $T_2$, but the conflicting write to $X$
in $T_2$, and the subsequent operations (due to the monotonicity
requirement of {\sc mav}) witness $T_1$. Thus, $T_2$ is aware of $T_1$
when it commits. Lastly, \iso{Serializable} isolation extends
$\mathtt{SnapshotSER}$ to also cover non-conflicting transactions that
write to variables read in the current transaction.
Fig.~\ref{fig:ansi-iso-eg-ser} shows an execution of an {\sc ser}
transacton $T_3$ and a {\sc mav transaction} $T_4$, which write to $X$
and $Y$, respectively. The snapshot witnessed by the {\sc ser}
transaction $T_3$ does not include $T_4$, but the write to $Y$ in
$T_4$, although non-conflicting, witnesses $T_1$ because $Y$ is read
by $T_1$. Visibility includes other operations of $T_4$ because of
{\sc mav}.
